---
title: "day79"
author: "csiu"
date: "May 14, 2017"
output:
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Still learning Scala, but today I thought I will show something new.


```{r message=FALSE}
library(dplyr)
library(stringr)
library(tidyr)
library(ggplot2)
```

## Data

http://data.vancouver.ca/datacatalogue/animalInventory.htm

```{r}
file_source <- "ftp://webftp.vancouver.ca/OpenData/csv/"
filenames <- c("DeadAnimals.csv", "LostAnimals.csv", "RegisteredAnimals.csv")

# for (f in filenames) {
#   download.file(url = paste0(file_source, f), destfile = paste0("data/", f))
# }
```

## Dead Animals

```{r message=FALSE}
dat <-
  sapply(
    sub("Animals.csv", "", filenames),
    function(a){
      readr::read_csv(paste0("data/", a, "Animals.csv"))
  })
```

Time frame:

```{r}
dat[["Dead"]]$DeadAnimalDate %>% range()
dat[["Dead"]]$DeadAnimalDate %>% {max(.) - min(.)}
```

```{r}
dat[["Dead"]] %>%
  {sprintf("Number of dead animals: %d", nrow(.))}
```

With 20 data points, it's easy to open the data in excel and look at it.
First thing I notice is that the majority (19) of dead animals are cats.

```{r}
ggtitle <- 
    "Domestic deceased pets picked up by Vancouver City staff"
subtitle <- sprintf(
  "%s day time frame (%s to %s); %s animals",
  dat[["Dead"]]$DeadAnimalDate %>% {max(.) - min(.)},
  dat[["Dead"]]$DeadAnimalDate %>% min,
  dat[["Dead"]]$DeadAnimalDate %>% max,
  dat[["Dead"]] %>% nrow
  )
cat_legend <- "
DSH = Domestic Short-hair
DLH = Domestic Long-hair
DMH = Domestic Medium-hair
NA = Dog
"

dat[["Dead"]] %>%
  mutate(
    cat = sub("Cat -[ ]?", "", CatOther),
    cat = sub("^(D.*?) -.*", "\\1", cat),
    cat = sub("Cat (D.*?) -.*", "\\1", cat)
    ) %>%
  group_by(cat) %>%
  summarise(count = n()) %>%
  ggplot(aes(x = reorder(cat, -count),
             y = count)) + 
  geom_bar(stat = "identity") +
  geom_text(y = 10.4, x = 1.7, label=cat_legend, vjust=1, hjust=0) + 
  xlab("Cat") +
  ylab("Dead") +
  labs(title = ggtitle,
       subtitle = subtitle)
```

## Animals that are reported lost

```{r}
lost_animals <-
  dat[["Lost"]] %>%
  filter(State == "Lost") %>%
  mutate(year = format(Date, "%Y") %>% as.integer())
```

```{r}
lost_animals %>%
  group_by(year) %>%
  summarise(count = n()) %>%
  arrange(year) %>%
  {
    ggplot(., aes(x = year, y = count)) +
      geom_point() +
      geom_line() +
      scale_x_continuous(breaks = seq(min(.$year), max(.$year), 1)) +
      theme(panel.grid.minor.x = element_blank()) +
      labs(title = "Animals reported lost per year",
           subtitle = "(to the Vancouver Animal Control Office)")
  }
ggsave("/Users/csiu/repo/blog/img/figure/2017-05-14/lost.png", width=7, height=5)
```

- Data for 2017 is not complete
- Interesting how there is a dip around 2008 and that no reports were made in 2008. Coincide with recession



```{r eval=FALSE, include=FALSE}
### Cleaning up the breed column

lost_animals %>%
  mutate(
    Breed = trimws(Breed, which = "both"),
    
    # Tidy breed
    Breed = sub("^\\?+$", "?", Breed),
    Breed = sub(".*Abyssinian.*|.*Abbysinian.*|.*Abithinian",
                "cat_Abyssinian", Breed),
    
    Breed = ifelse(grepl("parrot", Breed, ignore.case = T),
                   "parrot", Breed),
    # Breed = sub("Pitbull|Pit Bull|Pittbull|pitbull|Pitt Bull",
    #             "dog_pitbull", Breed),
    # Breed = gsub("[Ll]ab|Labrador",
    #             "dog_labrador", Breed),
    
    Breed = sub("husky", "dog", Breed, ignore.case = T),
    Breed = sub("pit[ ]?bull|Pittbull", "dog", Breed, ignore.case = T),
    Breed = sub("^Lab.*", "dog", Breed, ignore.case = T),
    Breed = sub("Chihuahua", "dog", Breed, ignore.case = T),
    Breed = sub("Shih Tzu|pug|Shiba Inu|Pomeranian", "dog", Breed, ignore.case = T),
    Breed = sub("^dog.*", "dog", Breed, ignore.case = T),
    Breed = sub(".*dog$", "dog", Breed, ignore.case = T),
    Breed = sub(".*pup.*", "dog", Breed, ignore.case = T),
    Breed = sub(".*pom*", "dog", Breed, ignore.case = T),
    
    Breed = sub("^cat.*", "cat", Breed, ignore.case = T),
    Breed = sub(".*cat$", "cat", Breed, ignore.case = T),
    Breed = sub(".*tabby*", "cat", Breed, ignore.case = T),
    Breed = ifelse(grepl("cat", Breed, ignore.case = T), "cat", Breed)
    

  ) %>% 
  
  
  group_by(Breed) %>%
  summarise(count = n()) %>%
  arrange(count) %>%
  #filter(grepl("cat", Breed, ignore.case = T)) %>%
  data.frame()

  # group_by(year, Breed) %>%
  # summarise(count = n()) %>%
  # {
  #   ggplot(., aes(x = year, y = count, color = Breed)) +
  #     geom_point() +
  #     geom_line() +
  #     scale_x_continuous(breaks = seq(min(.$year), max(.$year), 1)) +
  #     theme(panel.grid.minor.x = element_blank()) +
  #     labs(title = "Animals reported lost per year")
  # }

```
